Dynamics of Laterally Propagating Flames in X-ray Bursts

### 2019-05-13 xrb/paper.tex
more work

We work in 2D axisymmetric coordinates, but we advective a third
component of velocity, coming out of the simulation plane, that
participates in the Coriolis force (sometimes described as a 2.5D
simulation).  We will take the \castro\ $x$-coordinate to be the
cylindrical $r$ coordinate with corresponding velocity $u$, the
\castro\ $y$-coordinate to be the cylindrical $z$ coordinate with
corresponding velocity $v$, and the \castro\ $z$-coordinate to be the
cylindrical $\theta$ coordinate, with corresponding velocity $w$.  A
righthanded coordinate system has positive $w$ pointing out of the
page.  We will take $\Omegab = \Omega_0 \hat{\bf y}$ for the angular
rotation rate, and $\gb = -g \hat{\bf y}$ for the gravitational
acceleration, with $g$ constant.  With these choices, the Coriolis
force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omegab w \hat{x} - \Omegab u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref} that sets up via lateral
hydrostatic equilibrium.

@@

We use a general stellar equation of state with nuclei (treated as an
ideal gas), photons, and degenerate/relativistic electrons, as
described in \cite{timmes_swesty:2000}.  We use a simple reaction
network consisting of 4 nuclei, \isot{He}{4}, \isot{C}{12},
\isot{O}{16}, \isot{Fe}{56}.  Only the first 3 are evolved, with the
triple-$\alpha$ rate and $\isotm{C}{12}(\alpha, \gamma)\isotm{O}{16}$.
Screening of the reactions is also included.  The network is provided
in our open-source microphysics
routines\footnote{\url{https://github.com/starkiller-astro/Microphysics}}.

>>>>

Since the neutron star rotates, we work in a corotating frame, with
angular velocity $\Omegab$.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advective a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take the
\castro\ $x$-coordinate to be the cylindrical radial coordinate with
corresponding velocity $u$, the \castro\ $y$-coordinate to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
with corresponding velocity $w$.  A righthanded coordinate system has
positive $w$ pointing out of the page.  We will take $\Omegab =
\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omegab w \hat{x} - \Omegab u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref} that sets
up via lateral hydrostatic equilibrium.

@@

We use a general stellar equation of state with nuclei (treated as an
ideal gas), photons, and degenerate/relativistic electrons, as
described in \cite{timmes_swesty:2000}.  For reactions, we use a 13
isotope alpha chain, derived from the {\tt aprox13}
network~\citep{timmes_aprox13}.  For one of our runs, we use the
smaller 7-isotope network described in \citet{iso7}.  We integrate the
network using the VODE integration package~\citep{vode}, and out
implementation is provided in the StarKiller Microphysics
source~\citep{starkiller}.
